Computational hysteresis:\\ inter-task effects in human computation

### 2014-10-27 docs/publications/2014.09_SCIENCE/man.tex
more analysis work (dictionary) and manuscript work

\begin{mydef}
	\label{def:bias}
	{\upshape Maximum Bias:}
	Let $\mathcal{D}$ be a binary decision function, which takes a 
	work-product
	$r \in P \cup Q$ as input, and outputs a bit $\mathcal{D}(r)=b$, where 
	$b \in \{0,1\}$. Denote the \emph{bias} of $\mathcal{D}$, due to the
	different priming of $\mathcal{P}$ and $\mathcal{Q}$ by
	$\theta_\mathcal{D}(P,Q)$ and define it to be the difference 
	in the expected output of $\mathcal{D}$ when acting on work-products from 
	$P$ and $Q$:
	$$
	\theta_\mathcal{D}(P,Q) = 
		\left| 
			\mathrm{E}\left\{ \mathcal{D}(p) \right\} 
			- \mathrm{E}\left\{ \mathcal{D}(q) \right\} 
		\right|,
	$$
	where $p \in P$, $q \in Q$.  Then, define the difference in priming 
	between $\mathcal{P}$ and $\mathcal{Q}$ to be the worst-case bias, i.e. 
	the bias of the most susceptible decision:
	$$
	\theta(P,Q) = 
		\sup_\mathcal{D} \theta_\mathcal{D}(P,Q)
	$$
\end{mydef}
If $\mathcal{D}$ is the condition of an \texttt{if}-statement, then the 
definition above is the change in the fraction of the time that a given 
path of execution is followed due to HPU priming.

@@

\paragraph{Effects on label vocabulary.}
Inspecting the labels manually, one striking observation was that in
\textit{test-t.2}, workers that initially images of cultural scenes seemed
more likely to give generic labels, compared to those that initially labeled 
images depicting prepared meals.  Recall that, for these treatments, the test 
tasks contained images depicting prepared meals, which are very similar in
content to the initial tasks for the latter set of workers.  Workers that
had already labeled 5 images depicting food were less likely to attribute
the generic label ``food'' during the first test task, doing so only 16 times,
compared to 76 times by workers that had labeled culture-related images 
containing no food.  This observation suggests that, when workers label a 
series of very similar images.

One straightforward way to test whether workers from exp2.task.food are 
being more specific than those from exp2.taks.cult is to look at the number 
of unique labels they provided for each image.  Workers attributing more
specific labels should produce a larger vocabulary of unique labels for a 
given, which is indeed what we found (see Fig. 3B).  Workers from 
exp2.task.food produced as much as 33\% more unique labels per image than 
workers from exp2.task.cult, and the relative excess of the vocabulary size
followed the general trend of inter-task effects that we observed in 
Fig. 2B.

To confirm our expectation that the larger vocabulary corresponds to a 
the use of \textit{more nuanced} words, we used the 
hypernym-hyponym of wordnet to operationalize a notion of relative 
specificity.  We define the relative specificity of two sets of labels $P$ 
and $Q$ as:
$$
	S(P,Q) = \frac{1}{|P \times Q|}\sum_{p \in P} \sum_{q \in Q} 
		\left(\mathbf{1}_{[p>q]} - \mathbf{1}_{[q>p]}\right),
$$
where $\mathbf{1}_{[p>q]}$ evaluates to 1 if $p$ is a hyponym of $q$.
Using the above equation, we found that, for every images in the 
test set, the labels attributed by workers in exp2.task.food were more 
specific than those by workers in exp2.task.cult.

Thus, when workers perform a series of similar image labeling tasks it would
appear that they tend to avoid generic words, opting instead for more specific
ones, and yielding a richer vocabulary of labels.  Qualitative 
characterisation and classification is one of the most comman applications of 
microtask platforms, and this finding suggests that bundling similar tasks 
together substantially increases the specificity of HPU output.

Our observations of the effects on specificity would appear to be consistent
with the phenomenon of negative priming.
Negative priming occurs when a person becomes desensitized to non-salient 
stimuli to which she is repeatedly exposed. Consider that, workers that 
initially labeled images depicting prepared meals might no longer
regard the generic labels \texttt{food} or \texttt{meal} to be salient, and 
opt instead for \texttt{bread}, or \texttt{pasta}.  But for workers that
initially labeled cultural images, these rather generic labels might well seem
appropriate.

We are suggesting that, although workers are not instructed to compare images 
in any way, prior tasks nevertheless create a frame of reference relative to 
which later tasks are judged. This in turn influences the perception of 
salience. Thus, in a series of subjective characterization tasks that have 
very similar content, workers' focus will tend to be directed away from 
generic, shared attributes, towards those attributes that are specific and 
distinguishing.

To see how the initial tasks and the framing affected 
the focus of the workers' labels, we classified all the labels provided
on test tasks as either \texttt{culture}, \texttt{food}, \texttt{both}, or 
\texttt{neither} (we describe the label classification method in more detail
in the supplementary material).  Focusing on the labels provided to the first 
test task,
workers that had initially labeled cultural images provided far more cultural
labels during test tasks, and fewer food labels, than workers that initially
labeled images depicting ingredients (see \textbf{Fig. 2}).  On the other 
hand, the framing primes had no statistically significant effect.

\begin{table}
\centering
\begin{tabular}{ l  s s s s}

@@

Such a phenomenon would be consistent with the psychological mechanism known as
\textit{negative priming}.
\paragraph{The role of task position in inter-task effects.}
Up to this point, we have focused on the inter-task effects induced on 
the first test task.  Not suprisingly, this task was, in general, more 
strongly affected than the others.  It would stand to reason that, as workers 
proceed through the test images, priming induced by the initial tasks would
be ``washed out'', and replaced by the effects of the test tasks themselves,
which do not differ between treatments.

To investigate this, we performed a second experiment, in which we permuted
the test tasks in such a way that every test task appears in each of the
5 locations equally often.  This allows us to track the intertask effects for
a given test task as it is rotated through the five positions.  Workers were
again given one of two initial sets of tasks, which either involved labeling
utensils, cookware, and table settings (but no food), or isolated close-up
images of prepared meals.  We measured the performance of a Naive Bayes 
classifier in distinguishing whether labels attributed to a given test 
image were provided by a worker that labeled objects or meals, as shown in 
\textbf{Fig. 4}.

The strength of inter-task effects show a clear trend based on the position of
the test task.  We see that, in general, the first test image is most severely
affected, and the severity of effect drops off.  Curiously, the severity then
rises again for the second last, and, especially, final test image.  We 
speculate that this is related to the workers being aware that they are 
reaching the end of the batch of labeling tasks.  To be transparent about the
rate of compensation, we informed workers that they would complete a total
of ten image-labeling tasks.  Perhaps as the worker nears the end of the batch
of tasks, she reflects on the overall experience, and in activating the
memory of the early tasks, is partially re-primed.

>>>>

\begin{mydef}
	\label{def:bias}
	{\upshape Maximum Bias:}
	Let $\mathcal{D}$ be a binary decision function, which takes a 
	work-product
	$r \in P \cup Q$ as input, and outputs a bit $\mathcal{D}(r)=b$, where 
	$b \in \{0,1\}$. Denote the \emph{bias} of $\mathcal{D}$, due to the
	different priming of $\mathcal{P}$ and $\mathcal{Q}$ by
	$\theta_\mathcal{D}(P,Q)$ and define it to be the difference 
	in the expected output of $\mathcal{D}$ when acting on work-products from 
	$P$ and $Q$:
	$$
	\mathcal{D}(P,Q) = 
		\left| 
			\mathrm{E}\left\{ \mathcal{D}(p) \right\} 
			- \mathrm{E}\left\{ \mathcal{D}(q) \right\} 
		\right|,
	\theta_\mathcal{D}(P,Q) = 
		\left| 
			\mathrm{E}\left\{ \mathcal{D}(p) \right\} 
			- \mathrm{E}\left\{ \mathcal{D}(q) \right\} 
		\right|,
	$$
	where $p \in P$, $q \in Q$.  Then, define the difference in priming 
	between $\mathcal{P}$ and $\mathcal{Q}$ to be the worst-case bias, i.e. 
	the bias of the most susceptible decision:
	$$
	\theta(P,Q) = 
		\sup_\mathcal{D} \theta_\mathcal{D}(P,Q)
	$$
\end{mydef}
If $\mathcal{D}$ is the condition of an \texttt{if}-statement, then the 
definition above is the change in the fraction of the time that a given 
path of execution is followed due to HPU priming.

@@

\paragraph{Effects on label vocabulary.}
When manually inspecting the labels of \textit{test-t.2}, one striking 
observation was that workers that initially labeled images of cultural scenes 
were more likely to use generic words when labeling the test images, compared 
to workers that initially labeled images of meals.  For example, 76\% of the 
workers from \textit{exp2.task.cult} gave the very generic label 
\texttt{food}, compared to only 16\% of \textit{exp2.task.food} workers 
(recall that the test tasks for both treatments depicted prepared meals).
One interpretation is that workers who have already labeled 5 images relating
to food are more likely to focus on aspects of the images that are different
from one another, and refrain from so generic a term as \texttt{food}.

Based on this observation, we tested the degree of specificity of labels 
accross all treatments.  We appraoched measuring specificity in two ways.
Our first method was based on the reasoning that, if a given group of 
workers uses more specific labels, they will then provide, on the whole,
a larger set of unique labels for any given image.  For example,
suppose the workers from \textit{exp2.task.food} who do not use the label 
\texttt{food} had chosen more specific labels instead.  In being more 
specific, they must choose a particular detail to describe, from a large 
number of possibilities, and so are more likely to provide unique labels.
Indeed
for all images in the test set, workers from \textit{exp2.task.food} gave
more unique labels than did those from \textit{exp2.task.cult}, as shown 
in fig 2B.  Note how the the relative vocabulary sizes follows
the same trend, as a function of test task, as was observed for the 
classifier-based detection of general priming differences (c.f. fig X).

Our second approach to measuring relative specificity used the 
hypernym-hyponym relationships of wordnet.
We operationalized the relative specificity of two sets of labels $P$ 
and $Q$ as:
$$
	S(P,Q) = \frac{1}{|P \times Q|}\sum_{p \in P} \sum_{q \in Q} 
		\left(\mathbf{1}_{[p>q]} - \mathbf{1}_{[q>p]}\right),
$$
where $\mathbf{1}_{[p>q]}$ evaluates to 1 if $p$ is a hyponym of $q$.
Using the above equation, we found that, for every images in the 
test set, the labels attributed by workers in \textit{exp2.task.food} were 
more specific than those by workers in \textit{exp2.task.cult}.

Initialy we hypothesized that the similarity of initial and test tasks was
driving a relative increase in specificity.  While this seems reasonable,
we also notice a similar effect occuring for \textit{exp1.task} in which
the test tasks consisted of both food and non-food objects, and the initial
tasks were either food or non-food objects.  In this case, it is difficult
to argue that one set of initial tasks is more similar to another.  In fact
we devise a method to operationalize image similarity, and find that 
the initial images of \textit{exp1.task.food} are equally similar to the
test tasks as the initial images of \textit{exp1.task.obj}.  Furthermore,
we see strong effects on relative specificity for framing as well.

In general, it is of great practical interest to control the level of 
specificity of qualitative characeterzations.  In many applications the
task-designer will probably want obtain characterizations that are more
discriminative.  Thus, when designing microtasks, one should give careful
consideration to inter-task effects.  Unfortunately, we are not sure
why certain initial tasks produced increased specificity as opposed to others, 
and this is a question we intend to pursue in future research.

\section*{Supplementary Material}

\subsection*{Assessing task similarity}
\begin{table}
\centering
\begin{tabular}{ l  s s s s}

@@

Such a phenomenon would be consistent with the psychological mechanism known as
\textit{negative priming}.
